Article Title

### 2013-07-26 sections/Abstract.tex
introduction improvements

Correlations between the spiking activity of pairs of neurons are among the most familiar and useful descriptive statistics of neural activity. Multineuronal recordings provide richer information than the equivalent number of neuron pairs. For example, the full covariance matrix reveals correlated activity across the entire population as well as partial correlations between pairs.  Estimation of covariance matrices from finite recordings can be improved by regularization, \emph{i.e.}\;by imposing some kind of structure.  Optimal regularization must be determined empirically since its performance depends on how well the imposed structure captures the underlying regularities. For example, we can use low-rank parameterizations of the covariance matrix to account for common activity fluctuations across the recorded population. Conversely, if correlations are strongly influenced by a small fraction of possible pairwise linear associations between the observed neurons, we can impose sparsity on the inverse covariance matrix. We can also use a `sparse+low rank’ inverse covariance representation to account for both common fluctuations and pairwise interactions.

To select the optimal structure of neural correlations in a local neural circuit, we compared the performance of several covariance estimators on the activity of 100--300 neurons in mouse visual cortex: sample covariance, Ledoit-Wolf covariance shrinkage, factor analysis, sparse inverse covariance, and sparse+low-rank inverse covariance. We inferred instantaneous firing rates in 200 ms bins from the somatic calcium signals acquired with fast 3D random-access two-photon microscopy.  Each covariance estimator was optimized and evaluated by cross-validation. As expected, covariance shrinkage reliably outperformed the sample covariance estimate. In turn, factor analysis-based estimates significantly outperformed covariance shrinkage. Yet sparse inverse covariance with or without an additional low-rank component significantly outperformed both factor analysis and shrinkage estimators. The superior performance of the sparse inverse covariance estimator suggests the relative importance of detailed network interactions over common diffuse input in the circuit we studied.

>>>>

Correlations between the spiking activity of pairs of neurons are among the most familiar and useful descriptive statistics of neural activity. Multineuronal recordings provide richer information than the equivalent number of neuron pairs. For example, the full covariance matrix reveals correlated activity across the entire population as well as partial correlations between pairs.  Estimation of covariance matrices from finite recordings can be improved by regularization, \emph{i.e.}\;by imposing some kind of structure.  Optimal regularization must be determined empirically since its performance depends on how well the imposed structure matches the underlying regularities. For example, we can use low-rank parameterizations of the covariance matrix to account for common activity fluctuations across the recorded population. Conversely, if correlations are strongly influenced by a small fraction of possible pairwise linear associations between the observed neurons, we can impose sparsity on the inverse covariance matrix. We can also use a `sparse+low rank’ inverse covariance representation to account for both common fluctuations and pairwise interactions.

To select the optimal structure of neural correlations in a local neural circuit, we compared the performance of several covariance estimators on the activity of 100--300 neurons in mouse visual cortex: sample covariance, covariance shrinkage, factor analysis, sparse inverse covariance, and sparse+low-rank inverse covariance. We inferred instantaneous firing rates in 200 ms bins from the somatic calcium signals acquired with fast 3D random-access two-photon microscopy.  Each covariance estimator was optimized and evaluated by cross-validation. As expected, covariance shrinkage reliably outperformed the sample covariance estimate. In turn, factor analysis-based estimates significantly outperformed covariance shrinkage. Yet sparse inverse covariance with or without an additional low-rank component significantly outperformed both factor analysis and shrinkage estimators. The superior performance of the sparse inverse covariance estimator suggests the relative importance of detailed network interactions over common diffuse input in the circuit we studied.

### 2013-07-26 sections/Introduction.tex
introduction improvements

As the number of recorded cells increases, the usual estimations of parameters of the covariance matrix become increasingly noisy and biased. Numerical instabilities arise because, as the amount of recorded data increases only linearly, the number of free coefficients to be estimated in any parameterization of the covariance matrix increases quadratically. 
For example, the large eigenvalues from the sample covariance matrix are biased upward while small eigenvalues are biased downward \citep{Hayes:1981}. Similarly, the coefficients of the inverse covariance matrix require larger sample sizes to be estimated accurately from high-dimensional data than low-dimensional data.

In this study, we pursue two goals: 
\begin{enumerate}[A.\;\;]
	\item Devise good estimators of neural covariance matrices in recordings from large populations of neurons.
	\item Aid interpretation of population activity  by isolating the most essential features or structure of the neural covariance structure.
\end{enumerate}

These two goals are inextricably connected and are best addressed together. Interpretation of neural correlations can be aided by approximating the measured covariance matrix by a simplified, reduced form.  For example, \cite{Malmersjo:2013} analyze the spatial distribution and the graph-theoretical measures of only the highest, most significant correlations.

Estimation of parameters of the covariance matrix from data is inherently challenging, increasingly so for very high-dimensional data. The efficiency of an estimator relates to the rate of its convergence to the true value of the covariance matrix as the function of sample size. The usual estimator, the sample covariance matrix, has been shown to be substantially less efficient than a number of other estimators.  The efficiency of an estimator can be improved through \emph{regularization}, i.e.\;the deliberate biasing of the estimate toward a more simplified approximate representation \citep{Bickel:2006,Bickel:2008,Ledoit:2004,Schafer:2005}. A common intuitive explanation for the effect of regularization is that of adding some prior knowledge about the likely forms of the covariance matrix. Curiously, though, regularization need not reflect any plausible knowledge of the nature of covariance matrices in general or in the particular domain. Convergence can improve due to a purely statistical effect, sometimes known as \emph{Stein's phenomenon}, \emph{i.e.}~the favorable tradeoff between bias and variability, which arises \emph{anytime} a moderate bias toward a less variable target is mixed into an unbiased estimate.  Therefore, an estimator that is shown to converge faster on specific types of data does not necessarily contain within itself some wisdom of the statistical regularities in the data.  However, an estimator that does happen to accommodate the statistical structure of the data in a specific domain, it will gain additional advantage over an \emph{ad hoc} regularizer for that domain. Thus the optimal selection of the estimator for a specific problem with unknown statistical structure becomes an empirical question. And, if a clear winner emerges from a number of strong candidates, its regularization scheme can server as an effective statistical description of the covariance structure.

In this study, we compared the performance of several covariance estimators for spatially compact groups of 150--300 neurons in layers 2--4 in mouse primary visual cortex during visual stimulation.   The performance of the estimators was evaluated by computing the mean squared error between the optimized covariance estimate fitted to training data and the non-regularized covariance matrix from a separate test data set.  Low-rank covariance estimates performed significantly better than shrinkage estimators, but estimators with sparse partial correlations were more efficient still. Typically, between 3 and 16\% of neuronal pairs were connected by non-zero partial interactions.  Mixed sparse estimators with sparse partial correlations and low-rank components performed comparably.

As noted above, the optimal covariance estimator is domain-dependent. These specific findings may not spread to all neural circuits.
Improved accuracy estimation accuracy is not the main aim or the only benefit.  
The selection of the optimal regularization scheme in itself serves as an effective descriptor of the dominant correlation structures. The finding that identifiable subsets of partial correlations were important to describe the partial correlation makes such pairs as prime candidates as direct functional interactions.

>>>>

With large numbers of recorded cells, the usual estimations of the covariance matrix from finite recordings become ill-conditioned. Numerical instabilities arise because, as the amount of recorded data increases only linearly, the number of free coefficients to be estimated in any parameterization of the covariance matrix increases quadratically. 
For example, the large eigenvalues of large sample covariance matrices are biased upward while small eigenvalues are biased downward \citep{Hayes:1981}. Similarly, the coefficients of the inverse covariance matrix require larger sample sizes to be estimated accurately from high-dimensional data than low-dimensional data.

In this study, we pursue two goals: 
\begin{enumerate}[\qquad A.\;\;]
	\item Devise better estimators of neural covariance matrices in recordings from large populations of neurons.
	\item Aid interpretation of population activity  by isolating the essential structure of the neural covariance matrix.
\end{enumerate}

The two goals are inextricably connected as they both benefit from reduced, simplified approximations of the covariance matrix.

In estimation theory, the overall error of an estimate can be reduced by \emph{regularization}, which is the deliberate biasing of the solution toward a simplified, low-dimensional (`sparse') \emph{target estimate} \citep{Schafer:2005,Bickel:2006}.
Regularized estimators allow reaching a favorable tradeoff between bias and variance in order to minimize the total error.
Strikingly, some improvement can be produced even when the target estimate is chosen arbitrarily, which is sometimes described as ``Stein's phenomenon'' after its discoverer \cite{Stein:1956}.   However, when a sparse target estimate does happen to match the dominant features of the true covariance matrix, it will induce relatively little bias while retaining low variance.  For example, some covariance estimators have been motivated by a specific model of the underlying process such as single-factor models of the stock market, for example \citep{Ledoit:2003}. 
Since the true covariance structure is not known in practice, the performance of estimators is evaluated by how well their performance generalized to new data, outside the training sample by cross validation, for example.

The above consideration served as the basis of our approach.  We proposed several alternative candidate estimators of the neural covriance matrix in a given circuit, each motivated by a different hypothesis for the origin of neural correlations.  When a consistent winner emerges from cross-validated performance comparisons, the sparse estimate used in its regularization can be interpreted to reflect a simplified correlation structure supported by the emprical data.

We compared the performance of several covariance estimators for spatially compact groups of 150--300 neurons in layers 2--4 in mouse primary visual cortex during visual stimulation.   The performance of the estimators was evaluated by computing the mean squared error between the optimized covariance estimate fitted to training data and the non-regularized covariance matrix from a separate test data set.  Low-rank covariance estimates performed significantly better than shrinkage estimators, but estimators with sparse partial correlations were more efficient still. Typically, between 3 and 16\% of neuronal pairs were connected by non-zero partial interactions.  Mixed sparse estimators with sparse partial correlations and low-rank components performed comparably.
